------------------------------------------------------------------------

---
title: "Ximerakis workflow"
output: html_document
date: "2023-06-01"
---

## Seurat for Ximerakis

Adapted from [here](https://bioconductor.org/packages/release/bioc/vignettes/DropletUtils/inst/doc/DropletUtils.html#reading-in-10x-genomics-data) and [here](https://www.singlecellcourse.org/scrna-seq-analysis-with-bioconductor.html).

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DropletUtils")
library(DropletUtils)

BiocManager::install("Seurat")
library(Seurat)


library(dplyr)
library(stringr)

remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)
```

```{r}
#load Ximerakis paper FIXME this will need to be from group folder eventually, so amend paths at the start

# List of folder names
sample_names<- readLines("/home/s2268606/University_Directories/Project_23/External_datasets/Ximerakis/Ximerakis_BAM_names.txt")
#rename them for cell ranger counts folders
folder_names <- paste0("cr", sample_names)

# Loop through each folder
for (folder_name in folder_names) {
  folder_path=paste0("/home/s2268606/University_Directories/Project_23/External_datasets/Ximerakis/",folder_name, "/outs/raw_feature_bc_matrix/")
  current_data<-Read10X(data.dir =folder_path)
  
  #error trap for one of the variables... not sure which
  if (is.null(colnames(current_data))) {
    cat("Error: No cell names present in the input matrix for folder", folder_name, "\n")
    next  # Skip to the next iteration of the loop
  }
  
  # Assign the object to a variable with the folder name as the variable name
  assign(folder_name, current_data)
  
  # Make names for and create the Seurat objects
  objName <- paste0(folder_name, "obj")
  current_obj <- CreateSeuratObject(counts = current_data, project = "Ximerakis")
  assign(objName, current_obj)
}

#make the list of merging objects, omit the others with a shared name parts
matching_vars <- ls(pattern = 'obj')
matching_vars <- matching_vars[!matching_vars %in% c("current_obj", "objName")]

#merge the samples into one object
Ximerakis_total <- merge(crOX1Xobj, y= c(crOX2Xobj,crOX3Xobj,crOX4Xobj,crOX5Xobj,crOX6Xobj,crOX7Xobj,crOX8Xobj,crYX1Lobj,crYX2Lobj,crYX3Robj,crYX4Robj,crYX5Robj,crYX6Lobj,crYX7Robj,crYX8Lobj), add.cell.ids=c("OX1X", "OX2X", "OX3X", "OX4X", "OX5X", "OX6X", "OX7X", "OX8X", "YX1L", "YX2L","YX3R", "YX4R", "YX5R" ,"YX6L", "YX7R", "YX8L"), project="Ximerakis")

#cleaning up memory
# rm("crOX1X", "crOX2X", "crOX3X", "crOX4X", "crOX5X", "crOX6X", "crOX7X", "crOX8X", "crYX1L", "crYX2L","crYX3R", "crYX4R", "crYX5R" ,"crYX6L", "crYX7R", "crYX8L","matching_vars")
```

## Identify mitochondrial and ribosomal populations

```{r}
#identify mitochondrial and ribosomal protein gene portions, annotations syntax tailored to mouse
Ximerakis_total[["percent.mt"]]<- PercentageFeatureSet(Ximerakis_total, pattern = "^mt")
Ximerakis_total[["percent.rp"]]<- PercentageFeatureSet(Ximerakis_total, pattern = "^Rp")

#Generate genes per umi counts, from https://github.com/hbctraining/scRNA-seq/blob/master/lessons/04_SC_quality_control.md
Ximerakis_total$log10GenesPerUMI <- log10(Ximerakis_total$nFeature_RNA) / log10(Ximerakis_total$nCount_RNA)

Ximerakis_total$mitoRatio <- Ximerakis_total@meta.data$percent.mt / 100

```

## Initial data assessment

```{r}
VlnPlot(Ximerakis_total, features=c("nFeature_RNA","nCount_RNA","percent.mt", "percent.rp"), ncol=4, pt.size = 0.01)+ ggtitle("Preliminary Data assessment", subtitle= "Xiemrakis, unfiltered")
```
We can see a high proportion of mitochondrial reads which will need to be omitted, though this may be taken care of by other filters applied

### Add metadata, adding sample identity

adapted from <https://youtu.be/p49seH2_i8Y?t=312>

```{r}

#Check whether the whole block is needed here
sampleLabs<-names(Ximerakis_total@active.ident)

#Assigning names to the samples to keep identity
#From Chat GPT
sample_detect <- case_when(
  str_detect(sampleLabs, "OX1X") ~ "OX1X",
  str_detect(sampleLabs, "OX2X") ~ "OX2X",
  str_detect(sampleLabs, "OX3X") ~ "OX3X",
  str_detect(sampleLabs, "OX4X") ~ "OX4X",
  str_detect(sampleLabs, "OX5X") ~ "OX5X",
  str_detect(sampleLabs, "OX6X") ~ "OX6X",
  str_detect(sampleLabs, "OX7X") ~ "OX7X",
  str_detect(sampleLabs, "OX8X") ~ "OX8X",
  str_detect(sampleLabs, "YX1L") ~ "YX1L",
  str_detect(sampleLabs, "YX2L") ~ "YX2L",
  str_detect(sampleLabs, "YX3R") ~ "YX3R",
  str_detect(sampleLabs, "YX4R") ~ "YX4R",
  str_detect(sampleLabs, "YX5R") ~ "YX5R",
  str_detect(sampleLabs, "YX6L") ~ "YX6L",
  str_detect(sampleLabs, "YX7R") ~ "YX7R",
  str_detect(sampleLabs, "YX8L") ~ "YX8L",
  TRUE ~ NA_character_
)
#
Ximerakis_total@meta.data$sample<-sample_detect
Idents(object=Ximerakis_total)<-"sample"
rm(sampleLabs, sample_detect)
```


```{r}
#Save the data object
save(Ximerakis_total, file="./Rdata/Ximerakis_total_seurat.RData")
```


# To test

```{r}
# Rename columns from https://github.com/hbctraining/scRNA-seq/blob/master/lessons/04_SC_quality_control.md

#Create metadata variable to edit it without affecting core data
metadata_test<-Ximerakis_total@meta.data

#Saving cell barcodes as metainfo
metadata_test$cells<-rownames(metadata_test)

metadata_test <- metadata_test %>%
    dplyr::rename(nUMI = nCount_RNA,
                  nGene = nFeature_RNA)
```

### Counts per sample

This shows over 300k cells for some samples, indicating that minimum metrics must be set to remove junk.

```{r}
# Visualize the number of cell counts per sample with no filtering at all
metadata_test %>% 
  	ggplot(aes(x=sample, fill=sample)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Cell counts per sample", subtitle="Ximerakis, unfiltered")
```

## Setting a minimum gene count for quality control

20 genes minimum, still gives 80k cells in some samples!

```{r}
#Setting minimum feature count to filter cells with no genes UMIs vs. genes detected

metadata_test <- subset(Ximerakis_total, subset= nFeature_RNA > 20)@meta.data
```

```{r}
# Visualize the number of cell counts per sample with a minimum gene count set
metadata_test %>% 
  	ggplot(aes(x=sample, fill=sample)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Cell counts per sample", subtitle="Ximerakis, filtered to minimum 20 genes")
```

### UMI/ transcripts per cell

With line at 500 for ideal minimum

```{r}
# Visualize the number UMIs/transcripts per cell
metadata_test %>% 
  	ggplot(aes(color=sample, x=nUMI, fill= sample)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500)+
    ggtitle("UMI/transcripts per cell", subtitle="Ximerakis, filtered to minimum 20 genes")
```

### Genes per cell

```{r}
# Visualize the distribution of genes detected per cell via histogram
metadata_test %>% 
  	ggplot(aes(color=sample, x=nGene, fill= sample)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300)+
    ggtitle("Distribution of genes per cell", subtitle="Ximerakis, filtered to minimum 20 genes")

# Visualize the distribution of genes detected per cell via boxplot
metadata_test %>% 
  	ggplot(aes(x=sample, y=log10(nGene), fill=sample)) + 
  	geom_boxplot() + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Distribution of genes per cell", subtitle="Ximerakis, filtered to minimum 20 genes")
```

### UMI vs. genes detected

We want cells in top right

```{r,fig.dim = c(8, 6)}
#Some code commented out because it introduced infitine values with a log transform
# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs
metadata_test %>% 
  	ggplot(aes(x=nUMI, y=nGene, color=mitoRatio)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	#stat_smooth(method=lm) +
  	scale_x_log10() +
  	scale_y_log10() +
  	theme_classic() +
  	geom_vline(xintercept = 500) +
  	geom_hline(yintercept = 250) +
  	facet_wrap(~sample) +
    ggtitle("UMIs vs. genes detected", subtitle = "Ximerakis, minimum 20 genes")
```

### MitoRatio

We want data lower than 0.2

```{r}
# Visualize the distribution of mitochondrial gene expression detected per cell
metadata_test %>% 
  	ggplot(aes(color=sample, x=mitoRatio, fill=sample)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 0.2)+
    ggtitle("Distribution of mitochondrial gene expression per cell", subtitle="Ximerakis, filtered to minimum 20 genes")
```

```{r}
FeatureScatter((subset(Ximerakis_total, subset= nFeature_RNA > 20)), feature1 = "nFeature_RNA", feature2 = "percent.mt")
```

### Complexity

Should be above 0.8.

```{r}
# Visualize the overall complexity of the gene expression by visualizing the genes detected per UMI
metadata_test %>%
  	ggplot(aes(x=log10GenesPerUMI, color = sample, fill=sample)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = 0.8)+
    ggtitle("Complexity of gene expression (genes per UMI)", subtitle="Ximerakis, filtered to minimum 20 genes")
```

### Filtering metadata

```{r}
metadata_test<- metadata_test %>%
    filter(nUMI > 0, nGene >300, percent.mt <20)
```


## Repeat the QC analysis


```{r}
# Visualize the number of cell counts per sample with a minimum gene count set
metadata_test %>% 
  	ggplot(aes(x=sample, fill=sample)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Cell counts per sample", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")

table(metadata_test$sample)
```

### UMI/ transcripts per cell

With line at 500 for ideal minimum

```{r}
# Visualize the number UMIs/transcripts per cell
metadata_test %>% 
  	ggplot(aes(color=sample, x=nUMI, fill= sample)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500)+
    ggtitle("UMI/transcripts per cell", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")
```

### Genes per cell

```{r}
# Visualize the distribution of genes detected per cell via histogram
metadata_test %>% 
  	ggplot(aes(color=sample, x=nGene, fill= sample)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300)+
    ggtitle("Distribution of genes per cell", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")

# Visualize the distribution of genes detected per cell via boxplot
metadata_test %>% 
  	ggplot(aes(x=sample, y=log10(nGene), fill=sample)) + 
  	geom_boxplot() + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Distribution of genes per cell", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")
```

### UMI vs. genes detected

We want cells in top right

```{r,fig.dim = c(8, 6)}
#Some code commented out because it introduced infitine values with a log transform
# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs
metadata_test %>% 
  	ggplot(aes(x=nUMI, y=nGene, color=mitoRatio)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	#stat_smooth(method=lm) +
  	scale_x_log10() +
  	scale_y_log10() +
  	theme_classic() +
  	geom_vline(xintercept = 500) +
  	geom_hline(yintercept = 250) +
  	facet_wrap(~sample) +
    ggtitle("UMIs vs. genes detected", subtitle = "Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")
```

### MitoRatio

We want data lower than 0.2

```{r}
# Visualize the distribution of mitochondrial gene expression detected per cell
metadata_test %>% 
  	ggplot(aes(color=sample, x=mitoRatio, fill=sample)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 0.2)+
    ggtitle("Distribution of mitochondrial gene expression per cell", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")
```


### Complexity

Should be above 0.8.

```{r}
# Visualize the overall complexity of the gene expression by visualizing the genes detected per UMI
metadata_test %>%
  	ggplot(aes(x=log10GenesPerUMI, color = sample, fill=sample)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = 0.8)+
    ggtitle("Complexity of gene expression (genes per UMI)", subtitle="Ximerakis, filtered to minimum 0 transcripts, 300 genes and maximum 20% mitochondrial gene content")
```



### Assign filters to data

```{r}
subset
```

### Saving object 

```{r}
# Create .RData object to load at any time
save(metadata_test, file="Rdata/Ximerakis_filtered.RData")

```


# To keep
```{r}
#plot(density(Ximerakis_total@meta.data$nFeature_RNA))
VlnPlot(Ximerakis_total, features=c("nFeature_RNA","nCount_RNA","percent.mt", "percent.rp"), ncol=4, pt.size = 0.05)
```

## Plot each of these respectively, showing trends in cell health

```{r}
#show mitochondrial proportion in reads
FeatureScatter(Ximerakis_total, feature1 = "nCount_RNA", feature2 = "percent.mt")+ggtitle("UMIs per cell vs. % mitochondrial genes detected")

#show ribosomal proportion in reads
FeatureScatter(Ximerakis_total, feature1 = "nCount_RNA", feature2 = "percent.rp")+ggtitle("UMIs per cell vs. % ribosomal protein  genes detected")

#Ribosomal vs mitochondrial 
FeatureScatter(Ximerakis_total, feature1 = "percent.rp", feature2 = "percent.mt")+ggtitle("% ribosomal protein  genes detected vs. % mitochondiral genes detected")
```

There appears to be a shelf around 35% mt, indicating cell death. It seems that a couple of the Old mice have low mt, whilst a couple of the Young mice in pink have around 45%.

```{r}
#Show overall reads and gene counts for unfiltered data.
FeatureScatter(Ximerakis_total, feature1 = "nCount_RNA", feature2="nFeature_RNA",raster=FALSE, pt.size = 0.05)+ geom_smooth(method="lm")+ ggtitle("UMIs per cell vs. number of genes ")
```

Seems normal, though values are lower than our data

## Filter anything with more than 20% mito content

Creating new data variable for this. Results in the filtering of 29% of cells to leave 1,553,556 cells. Biologically, filtering

```{r}
#Create data with less than 20 % mt
Ximerakis_mitofilter20<-subset(Ximerakis_total,subset= percent.mt <20)

FeatureScatter(Ximerakis_mitofilter20, feature1 = "nCount_RNA", feature2 = "percent.mt")+ggtitle("UMIs per cell vs. % mitochondrial genes detected (filtered by mt%)")

#compared to the first assessment
VlnPlot(Ximerakis_mitofilter20, features=c("nFeature_RNA","nCount_RNA","percent.mt", "percent.rp"), ncol=4, pt.size = 0.05)+ggtitle("Mitochondria-fitlered summary")
```

## Filter out anything with less than 20 reads

This removes around 3/4 of data. Not filtered by mitochondria, just reads. Results in the filtering of 29% of cells to leave 1,035,865 cells ??

```{r}
Ximerakis_readfilter<-subset(Ximerakis_total,subset= nCount_RNA > 20)
#Verify the effects
FeatureScatter(Ximerakis_readfilter, feature1 = "nCount_RNA", feature2 = "percent.mt")+ggtitle("UMIs per cell vs. % mitochondrial genes detected with read filtering")
#compared violin plot to the first assessment
VlnPlot(Ximerakis_readfilter, features=c("nFeature_RNA","nCount_RNA","percent.mt", "percent.rp"), ncol=4, pt.size = 0.05)+ggtitle("Read-fitlered summary")
```

## Applying all filters

Filtering by at elast 10 genes, less than 20% mitochondrial gene content and more than 20 UMIs. Would give 390,282 cells.

```{r}
barplot(table(Idents((subset(Ximerakis_total, subset= nFeature_RNA >10 & percent.mt <20 & nCount_RNA > 20)))))
```

## Assign a 'pass' status for these QC filters

```{r}
#Doesn't work yet

#At least 250 genes
Ximerakis_total[['QC']] <- ifelse(Ximerakis_total@meta.data$nFeature_RNA < 250,'Low_nFeature',Ximerakis_total@meta.data$QC)


Ximerakis_total[['QC']] <- ifelse(Ximerakis_total@meta.data$nFeature_RNA < 250 & Ximerakis_total@meta.data$QC != 'Pass' &Ximerakis_total@meta.data$QC != 'Low_nFeature',paste('Low_nFeature',Ximerakis_total@meta.data$QC,sep = ','),Ximerakis_total@meta.data$QC)

#Less than 20% mitochondrial
Ximerakis_total[['QC']] <- ifelse(Ximerakis_total@meta.data$percent.mt > 20 & Ximerakis_total@meta.data$QC =='Pass','High_MT',Ximerakis_total@meta.data$QC)

Ximerakis_total[['QC']] <- ifelse(Ximerakis_total@meta.data$nFeature_RNA < 250 & Ximerakis_total@meta.data$QC != 'Pass' &Ximerakis_total@meta.data$QC != 'High_MT',paste('High_MT',Ximerakis_total@meta.data$QC,sep = ','),Ximerakis_total@meta.data$QC)

table(Ximerakis_total[['QC']])
```

```{r}
#filter out the cells with mitochondrial genes more than 5% and genes with more than .... genes as instructed by Nick
Ximerakis_total<- subset(Ximerakis_total, 
                         subset= percent.mt <20 &
                           #FIXME add more filters here
                         )
#                         & nFeature_RNA > 1000)      ##FIXME filter needs adjusting
```

## Doublet finder for all samples

```{r}
#adapted from https://youtu.be/p49seH2_i8Y , https://github.com/kpatel427/YouTubeTutorials/blob/5b3c795cecafb22f0e992da069eae2efd6b5cb95/singleCell_doublets.R , https://rpubs.com/kenneditodd/doublet_finder_example 

Ximerakis.list <- SplitObject(Ximerakis_total, split.by = "sample")

# Perform doublet detection on each sample using lapply
Ximerakis.list <- lapply(Ximerakis.list, function(sample) {
  print(paste0("Sample ",sample@meta.data$sample[1]))
  # Pre-process seurat object with standard seurat workflow
  sample <- NormalizeData(sample)
  sample <- SCTransform(sample)
  sample <- RunPCA(sample)
  
  # Find significant PCs
  stdv <- sample[["pca"]]@stdev
  sum.stdv <- sum(sample[["pca"]]@stdev)
  percent.stdv <- (stdv / sum.stdv) * 100
  cumulative <- cumsum(percent.stdv)
  co1 <- which(cumulative > 90 & percent.stdv < 5)[1]
  co2 <- sort(which((percent.stdv[1:length(percent.stdv) - 1] -
                       percent.stdv[2:length(percent.stdv)]) > 0.1),
              decreasing = TRUE)[1] + 1
  min.pc <- min(co1, co2)
  
  # Finish pre-processing
  sample <- RunUMAP(sample, dims = 1:min.pc)
  sample <- FindNeighbors(object = sample, dims = 1:min.pc)              
  sample <- FindClusters(object = sample, resolution = 0.1)
  
  # pK identification (no ground-truth)
  sweep.list <- paramSweep_v3(sample, PCs = 1:min.pc, sct = TRUE)
  sweep.stats <- summarizeSweep(sweep.list)
  bcmvn <- find.pK(sweep.stats)
  
  # Optimal pK is the max of the bomodality coefficient (BCmvn) distribution
  bcmvn.max <- bcmvn[which.max(bcmvn$BCmetric),]
  optimal.pk <- as.numeric(bcmvn.max$pK)

  # Homotypic doublet proportion estimate
  annotations <- sample@meta.data$seurat_clusters
  homotypic.prop <- modelHomotypic(annotations) 
  nExp.poi <- round(0.087 * nrow(sample@meta.data))
  nExp.poi.adj <- round(nExp.poi * (1 - homotypic.prop))
  
  # Run DoubletFinder
  sample <- doubletFinder_v3(seu = sample, 
                             PCs = 1:min.pc, 
                             pK = optimal.pk,
                             nExp = nExp.poi.adj, 
                             sct = TRUE, 
                             pN=0.25)
  
  return(sample)
  
})

#Rename the columns
Ximerakis.list <- lapply(Ximerakis.list, function(sample) {
  doublet_col_index <- which(grepl("DF.classifications_", colnames(sample@meta.data)))
  sample$doublet_finder <- sample@meta.data[doublet_col_index]
  sample@meta.data[doublet_col_index]<-NULL
  return(sample)
})

#subset data by singlet status
Ximerakis.list <- lapply(Ximerakis.list, function(sample) {
  sample<- subset(sample, subset= doublet_finder == 'Singlet')
  return(sample)
  })
```

```{r}
#Show doublet info 
FeatureScatter(Ximerakis_total, feature1 = "nFeature_RNA", feature2 = "Doublet_score")

```

```{r}
#Misc code
# code from https://github.com/kpatel427/YouTubeTutorials/blob/5b3c795cecafb22f0e992da069eae2efd6b5cb95/singleCell_doublets.R

#show the capture of variability in the PCs
ElbowPlot(Ximerakis.list[[1]])+ggtitle("Variability captured at each level of PC in PCA (sample OX1X)")

# # visualize doublets
# DimPlot(x, reduction = 'umap', group.by = "DF.classifications_0.25_0.21_691")


# # number of singlets and doublets
# table(x@meta.data$DF.classifications_0.25_0.21_691)

  #FIXME is this overwriting to only have singlets?
  # # Subset and return singlets
  # singlets <- subset(sample, doublet_finder == "Singlet")
  # return(singlets)  
# # Merge singlets from all samples
# Ximerakis.singlets <- merge(list items)
```
